network_analysis

### 2020-09-21 analyses/network_analysis_v2.Rmd
update prior to trying to recover from a broken projecy

  for(i in 1:iter) {
    temp <- net %>%
      semi_join(t_filter) %>%
      dplyr::select(site, date, year) %>%
      unique() %>%
      group_by(site, year) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      unite(webID, c(site, year), sep = "_") %>%
      ungroup() %>%
      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      #map(specieslevel, level = "higher", index = "d") %>%
      map(specieslevel, level = "higher", index = c("d", "resource range")) %>%
      map(rownames_to_column) %>%
      map(as_tibble) %>%
      bind_rows(.id = "site_year") %>% # Collapse list into big data frame
      rename(bb.sp = rowname) %>%
      mutate(iteration = rep(i, n()))

@@

  for(i in 1:iter) {
    temp <- net %>%
      semi_join(t_filter) %>%
      dplyr::select(site, date, year) %>%
      unique() %>%
      group_by(site, year) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      group_by(site, year) %>%
      unite(webID, c(site, year), sep = "_") %>%
      #filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
      ungroup() %>%
      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(networklevel, level = "higher", weighted = TRUE, index = c("niche overlap")) %>% # We probably care mainly about the BB level of the network
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(site_year = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n()))

@@

site_BB_aux.sp <- net %>%
  group_by(site, date, bb.sp) %>% # group by bb.sp*transect (and other variables we want to retain)
  summarize(bb.abund = n()) %>% # get bb.sp*transect abundance
  ungroup %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>%
  mutate(year = factor(year(date))) %>%
  group_by(site, year, bb.sp) %>%
  summarize(bb.sp.abund = mean(bb.abund))

@@

wma1_2010 <- filter(net, site == "wma1", year == "2010")

@@

spmet_mean <- spmet %>%
  group_by(site, year) %>%
  #summarize(d = weighted.mean(d, bb.sp.abund, na.rm = TRUE)) %>%
  summarize(resource.range = weighted.mean(resource.range, bb.sp.abund, na.rm = TRUE)) %>%
  ungroup()

### Merge network metrics and auxiliary data
netmet <- bind_rows(netmet_boot_2012, 
                    netmet_boot_2011,
                    netmet_boot_2010) %>%
  ungroup() %>%
  spread(metric, value) %>%
  left_join(spmet_mean, by = c("site", "year")) %>%
  left_join(site_BB_aux, by = c("site", "year")) %>%
  inner_join(site_FL_aux, by = c("site", "year")) %>% # one site-year is missing from floral survey data
  left_join(site_data, by = "site") %>%
  mutate(year = factor(year, levels = c("2010", "2011", "2012")),
         site = factor(site),
         management = factor(management))
```

@@

### Elevation-focused modeling
```{r}
#### Niche overlap
gam_niche.overlap.1 <- gam(niche.overlap.HL ~ 
                             s(year, bs = "re") + 
                             s(bb.transects) + 
                             s(elev.mean, by = year, k = 10),
                           select = TRUE,
                           data = netmet,
                           method = "REML") %>% getViz()

@@

# bb.transects was zeroed out by select = TRUE, so this plot should reflect the model above perfectly
ggplot(netmet, aes(elev.mean, niche.overlap.HL)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 10)) +
  facet_wrap(~year)

##### Community mean d'
gam_mean_d.1 <- gam(d ~ 
                      s(year, bs = "re") + 
                      s(bb.transects) +
                      s(elev.mean, by = year, k = 10),
                    select = TRUE,
                    data = netmet,
                    method = "REML") %>% getViz()

check.gamViz(gam_mean_d.1)
concurvity(gam_mean_d.1, full = TRUE)
concurvity(gam_mean_d.1, full = FALSE)
summary(gam_mean_d.1)
print(plot(gam_mean_d.1, allTerms = TRUE), pages = 1)

ggplot(netmet, aes(elev.mean, d)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 10)) +
  facet_wrap(~year)

ggplot(netmet, aes(bb.transects, d)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~year, scales = "free_x")

# I would be reluctant to interpret the year effect biologically; more likely an artifact of sampling intensity
ggplot(netmet, aes(year, d)) +
  geom_boxplot()

# It is interesting that, while elevation explains a lot of niche overlap and virtually none of specialization, niche overlap and specialization are strongly correlated with each other.
ggplot(netmet, aes(niche.overlap.HL, d)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~year)

##### Resource range
gam_mean_rr.1 <- gam(resource.range ~ 
                      s(year, bs = "re") + 
                      s(bb.transects) +
                      s(elev.mean, by = year, k = 10),
                    select = TRUE,
                    data = netmet,
                    method = "REML") %>% getViz()

@@

## Species-level resource range
gam_rr.sp_I <- gam(resource.range ~ 
                    s(year, bs = "re") + 
                    #s(site, bs = "re") +
                    #management +
                    s(bb.transects) +
                    s(bb.sp, bs = "re") +
                    s(elev.mean, by = bb.sp, bs = "tp", k = 5),
                  #s(fl.rich, k = 8) +
                  #s(log.fl.abund, k = 8) +
                  #s(bb.sp.abund, k = 5),
                  select = TRUE,
                  data = spmet,
                  method = "REML") %>% getViz()

@@

@@

#### Community mean d'
gam_mean_d.2 <- gam(d ~ 
                      s(year, bs = "re") + 
                      s(management, bs = "re") +
                      s(bb.transects) +
                      s(elev.mean, by = year, k = 10) +
                      s(bb.rich, k = 5) +
                      s(log.fl.abund, k = 5),
                    select = TRUE,
                    data = netmet,
                    method = "REML") %>% getViz()

check.gamViz(gam_mean_d.2)
concurvity(gam_mean_d.2, full = TRUE)
concurvity(gam_mean_d.2, full = FALSE)
summary(gam_mean_d.2)
print(plot(gam_mean_d.2, allTerms = TRUE), pages = 1)

ggplot(netmet, aes(bb.rich, d)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~year)

ggplot(netmet, aes(log.bb.abund, d)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~year)

ggplot(netmet, aes(log.fl.abund, d)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~year)

ggplot(netmet, aes(fl.rich, d)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~year)
```

@@

>>>>

  for(i in 1:iter) {
    temp <- net %>%
      semi_join(t_filter) %>%
      dplyr::select(site, date, year) %>%
      unique() %>%
      group_by(site, year) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      unite(webID, c(site, year), sep = "_") %>%
      ungroup() %>%
      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      #map(specieslevel, level = "higher", index = "d") %>%
      map(specieslevel, level = "higher", index = c("d", "resource range", "degree", "partner diversity")) %>%
      map(rownames_to_column) %>%
      map(as_tibble) %>%
      bind_rows(.id = "site_year") %>% # Collapse list into big data frame
      rename(bb.sp = rowname) %>%
      mutate(iteration = rep(i, n()))

@@

  for(i in 1:iter) {
    temp <- net %>%
      semi_join(t_filter) %>%
      dplyr::select(site, date, year) %>%
      unique() %>%
      group_by(site, year) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      group_by(site, year) %>%
      unite(webID, c(site, year), sep = "_") %>%
      #filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
      ungroup() %>%
      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(networklevel, level = "higher", weighted = TRUE, index = c("niche overlap", "web asymmetry", "weighted NODF", "partner diversity", "C score")) %>% # We probably care mainly about the BB level of the network
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(site_year = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n()))

@@

site_BB_aux.sp <- net %>%
  group_by(site, date, bb.sp) %>% # group by bb.sp*transect (and other variables we want to retain)
  summarize(bb.abund = n()) %>% # get bb.sp*transect abundance
  ungroup %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>%
  mutate(year = factor(year(date))) %>%
  group_by(site, year, bb.sp) %>%
  summarize(bb.sp.abund = mean(bb.abund),
            log.bb.abund = log(mean(bb.abund)),
            bb.transects = length(unique(date)))

@@

@@

spmet_mean <- spmet %>%
  group_by(site, year) %>%
  summarize(resource.range = weighted.mean(resource.range, bb.sp.abund, na.rm = TRUE),
            d = weighted.mean(d, bb.sp.abund, na.rm = TRUE),
            degree = weighted.mean(degree, bb.sp.abund, na.rm = TRUE),
            partner.diversity = weighted.mean(partner.diversity, bb.sp.abund, na.rm = TRUE)) %>%
  #summarize(resource.range = mean(resource.range, na.rm = TRUE)) %>%
  ungroup()

### Merge network metrics and auxiliary data
netmet <- bind_rows(netmet_boot_2012, 
                    netmet_boot_2011,
                    netmet_boot_2010) %>%
  ungroup() %>%
  spread(metric, value) %>%
  left_join(spmet_mean, by = c("site", "year")) %>%
  left_join(site_BB_aux, by = c("site", "year")) %>%
  inner_join(site_FL_aux, by = c("site", "year")) %>% # one site-year is missing from floral survey data
  left_join(site_data, by = "site") %>%
  mutate(year = factor(year, levels = c("2010", "2011", "2012")),
         site = factor(site),
         management = factor(management),
         threshold = factor(case_when(
           elev.mean > 1300 ~ "above",
           elev.mean <= 1300 ~ "below"
           ))
         )
```

@@

### Elevation-focused modeling
```{r}
ggplot(netmet, aes(elev.mean, C.score.HL,)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

#### Niche overlap
gam_niche.overlap.1 <- gam(niche.overlap.HL ~ 
                             s(year, bs = "re") + 
                             s(bb.transects) + 
                             s(log.bb.abund) +
                             #s(bb.rich) +
                             s(elev.mean, by = year, k = 10),
                           select = TRUE,
                           data = netmet,
                           method = "REML") %>% getViz()

@@

ggplot(netmet, aes(elev.mean, niche.overlap.HL)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 10)) +
  facet_wrap(~year)

ggplot(netmet, aes(log.bb.abund, niche.overlap.HL)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 10)) +
  facet_wrap(~year)

ggplot(netmet, aes(bb.rich, niche.overlap.HL)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 10)) +
  facet_wrap(~year)

#### Niche overlap
gam_C.score.1 <- gam(C.score.HL ~ 
                             s(year, bs = "re") + 
                             s(bb.transects) + 
                             s(log.bb.abund) +
                             #s(bb.rich) +
                             s(elev.mean, by = year, k = 10),
                           select = TRUE,
                           data = netmet,
                           method = "REML") %>% getViz()

check.gamViz(gam_C.score.1)
concurvity(gam_C.score.1, full = TRUE)
concurvity(gam_C.score.1, full = FALSE)
summary(gam_C.score.1)
print(plot(gam_C.score.1, allTerms = TRUE), pages = 1)

ggplot(netmet, aes(elev.mean, C.score.HL)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 10)) +
  facet_wrap(~year)

ggplot(netmet, aes(log.bb.abund, C.score.HL)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 10)) +
  facet_wrap(~year)

ggplot(netmet, aes(bb.rich, C.score.HL)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 10)) +
  facet_wrap(~year)

### NODF
gam_NODF.1 <- gam(weighted.NODF ~ 
                             s(year, bs = "re") + 
                             s(bb.transects) + 
                             s(log.bb.abund) +
                             #s(bb.rich) +
                             s(elev.mean, by = year, k = 10),
                           select = TRUE,
                           data = netmet,
                           method = "REML") %>% getViz()

check.gamViz(gam_NODF.1)
concurvity(gam_NODF.1, full = TRUE)
concurvity(gam_NODF.1, full = FALSE)
summary(gam_NODF.1)
print(plot(gam_NODF.1, allTerms = TRUE), pages = 1)

##### Resource range
gam_mean_rr.1 <- gam(resource.range ~ 
                      s(year, bs = "re") + 
                      s(bb.transects) +
                       s(log.bb.abund) +
                      s(elev.mean, by = year, k = 10),
                    select = TRUE,
                    data = netmet,
                    method = "REML") %>% getViz()

@@

## Species-level resource range
gam_rr.sp_I <- gam(resource.range ~ 
                    s(year, bs = "re") + 
                    #s(site, bs = "re") +
                    #management +
                    s(bb.transects) +
                    s(bb.sp, bs = "re") +
                     s(log.bb.abund, k = 5) +
                    s(elev.mean, by = bb.sp, bs = "tp", k = 5),
                  #s(fl.rich, k = 8) +
                  #s(log.fl.abund, k = 8) +
                  #s(bb.sp.abund, k = 5),
                  select = TRUE,
                  data = spmet,
                  method = "REML") %>% getViz()

@@

ggplot(spmet, aes(log.bb.abund, resource.range)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year, scales =  "free")

@@

#### Resource range
gam_mean_rr.2 <- gam(resource.range ~ 
                      s(year, bs = "re") + 
                      #s(management, bs = "re") +
                      s(bb.transects) +
                      s(elev.mean, by = year, k = 10) +
                      #s(bb.rich, k = 5) +
                       s(fl.rich, k = 5) +
                       s(log.bb.abund, k = 5),
                     # s(log.fl.abund, k = 5),
                    select = TRUE,
                    data = netmet,
                    method = "REML") %>% getViz()

check.gamViz(gam_mean_rr.2)
concurvity(gam_mean_rr.2, full = TRUE)
concurvity(gam_mean_rr.2, full = FALSE)
summary(gam_mean_rr.2)
print(plot(gam_mean_rr.2, allTerms = TRUE), pages = 1)

ggplot(netmet, aes(bb.rich, resource.range)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~year)

ggplot(netmet, aes(log.bb.abund, resource.range)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~year)

ggplot(netmet, aes(log.fl.abund, resource.range)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~year)

ggplot(netmet, aes(fl.rich, resource.range)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~year)
```

@@

## Threshold-based network analysis: do network properties differ meaningfully above and below the threshold (~1300m) evident in my abundance and beta-diversity analyses?
```{r}
